A 38 Million Year Old Mini-Neptune in the Kepler Field

### 2021-09-24 paper/ms.tex
update ms

Beyond the kinematic and rotational lines of evidence for the youth of
\sn, there is a final consistency check: the abundance of
Li\,\textsc{I} as inferred from the 6708\,\AA\ doublet \citep[{\it
e.g.},][]{soderblom_ages_2014}.  We acquired an iodine-free spectrum
from Keck/HIRES on the night of 2021 March 26 using the standard setup
and reduction techniques of the California Planet Survey
\citep{howard_cps_2010}.  Following the equivalent width measurement
procedure described by \citet{bouma_2021_ngc2516}, we find ${\rm
EW}_{\rm Li} = 233^{+5}_{-7}$\,m\AA.   %Note that this value does not
%correct for the Fe\,\textsc{I} blend at 6707.44\AA. 
Given the effective temperature of \sn\ (Table~\ref{tab:starparams}),
this measurement is in agreement with expectations for a 38\,Myr star
({\it e.g.}, as measured in IC~2602 by
\citealt{randich_gaiaeso_2018}).  It is also larger than any ${\rm
EW}_{\rm Li}$ found by \citet{berger_identifying_2018} in their
analysis of 1{,}301 Kepler spectra.

@@

		\vspace{-0.2cm}	
		\subfloat{
			\includegraphics[width=0.5\textwidth]{f5a.pdf}
			\includegraphics[width=0.5\textwidth]{f5b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The light curve of Kepler\,1627.}
    {\it Top}: 
    The Kepler data span 1{,}437 days (3.9 years), sampled at
    30 minute cadence;  a 100 day segment is shown.  The
    top panel shows the \texttt{PDCSAP} median-subtracted flux in
    units of parts-per-thousand ($\times 10^{-3}$).  The dominant
    signal is induced by starspots.  The stellar
    variability model (orange line) is subtracted below, revealing the
    transits of \pn.  The
    online Figure Set spans the entire 3.9 years of observations.
    {\it Bottom}:
    Phase-folded transit of Kepler 1627Ab with stellar variability
    removed.  Windows over 20 hours ({\it left}) and the entire orbit
    ({\it right}) are shown, and the residual after subtracting the
    transit in the bottom-most row.  The 2-$\sigma$ model uncertainties
    and the best-fit model are the light purple band and the dark
    purple line.  Gray points are individual flux measurements; black
    points bin these to 20 minute intervals, and have a representative
    1-$\sigma$ error bar in the lower right of each panel.  The
    asymmetric residual during transit appears to be robust to
    different detrending methods.
    \label{fig:lc}
  }
\end{figure*}

@@

The second line of confirmatory evidence comes by analyzing the
individual Kepler transit times. We isolated each of the 144 observed
transits to within $\pm4.5$\,hr of each transit, and fitted each
window with both {\it i)} a local second-order polynomial and transit,
and {\it ii)} a local linear trend.  We let the mid-time of each
transit float, and then calculated the residual between the measured
mid-time and that of a periodic orbit.  This residual, the transit
timing variation (TTV), is plotted in Figure~\ref{fig:obliquity}
against the local linear slope.  A significant correlation of $-0.0564
\pm 0.0098\ {\rm ppt\,day}^{-1}$ is observed.  Fewer than ten Kepler
Objects of Interest have shown this correlation
\citep{holczer_time_2015}, which is most readily interpreted as a TTV
induced by unresolved starspot crossings \citep{mazeh_time_2015}.
This is only possible if the planet transits the primary star, which
excludes a background eclipsing binary scenario, and therefore
confirms that \pn\ is a planet.  It also implies that the planet's
orbit is prograde.  The latter point assumes that the dominant
photometric variability is induced by dark spots, and not bright
faculae.  Given the observed transition of Sun-like stellar
variability from spot to faculae-dominated regimes between young and
old ages, we expect this assumption to be reasonably secure
\citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.

A third supporting line of evidence for the planetary interpretation
also exists, though we consider it less definitive than the stellar
density and TTV-local slope correlation.  We observed a transit of
\pn\ on the night of 2021 Aug 7 simultaneously with Keck/HIRES and
MuSCAT3.  We scheduled the observations using the ephemeris of
\citet{holczer_transit_2016}.  Although we did not detect the
Rossiter-McLaughlin (RM) anomaly, the multi-band MuSCAT3 light curves
show the transit (Figure~\ref{fig:obliquity}, lower panels).  Fitting
the MuSCAT3 photometry with a model that lets the transit depths vary
across each bandpass, we find {\it griz} depths consistent with the
Kepler depth at 0.6, 0.3, 0.3, and 1.1-$\sigma$ respectively.  The
MuSCAT3 observations also suggest a transit duration $17.4\pm3.6\ {\rm
minutes}$ shorter than the Kepler transits; further photometric
follow-up could help confirm whether the transit duration is indeed
changing.  For our RM analysis, the details are discussed in
Appendix~\ref{app:rm}.  While the velocities are marginally more
consistent with a prograde or polar orbit than a retrograde orbit, the
spot-corrected exposure-to-exposure scatter ($\sigma_{\rm RV}\approx
50$\ms) is significantly larger than the expected velocity anomaly
($\Delta v_{\rm RM}\approx 20\,{\rm m\,s}^{-1}$).  We are therefore
not in a position to claim a spectroscopic detection of the RM effect,
nor a precise quantification of the stellar obliquity.

@@

Kepler\,1627Ab provides a new extremum in the ages of the Kepler
planets, and opens multiple avenues for further study.  Further
observations of spectroscopic transits could yield a quantification of
the stellar obliquity, and would confirm or refute the expected
prograde orbital geometry.  Separately, transit spectroscopy aimed at
detecting atmospheric outflows could yield insight into the
evolutionary state of the atmosphere \citep[{\it
e.g.},][]{ehrenreich_giant_2015,spake_helium_2018,vissapragada_2020}.
Observations aimed at quantifying the amount of high-energy
irradiation incident on the planet would complement these efforts, by
helping to clarify the expected outflow rate \citep[{\it
e.g.},][]{poppenhaeger_2021}.  Finally, a challenging but informative
quantity to measure would be the planet's mass.  This might be
possible through a high-cadence multi-color radial velocity campaign,
or through long-term monitoring of the transit times and durations.
Measured at sufficient precision, the mass, combined with the known
age and size, would yield constraints on both the planet's composition
and its initial entropy \citep{owen_constraining_2020}.

More immediately, the Kepler data may yet contain additional
information.  For instance, one possible explanation for the transit
asymmetry shown in Figure~\ref{fig:lc} is that of a dusty asymmetric
outflow.  Other possible explanations such as favored longitudes for
starspot crossings seem strained for reasons discussed in
Appendix~\ref{app:gptransit}.  Dusty outflows are theoretically
expected for young mini-Neptunes, and the amplitude of the observed
asymmetry is also roughly consistent with predictions
\citep{wang_dai_2019}.  Beyond the planetary transits,
Appendix~\ref{app:flare} highlights a separate oddity in the Kepler
data, in the arrival time distribution of stellar flares.  We
encourage its exploration by investigators more versed in the topic
than ourselves.

In the context of the transiting planet population, \pn\ is among the
youngest known (Figure~\ref{fig:rp_period_age}).  Comparable systems
with precise ages include K2-33, DS~Tuc, HIP~67522,
TOI~837, the two-planet AU~Mic system, and the four-planet V1298~Tau
system
\citep{Mann_K2_33b_2016,David_et_al_2017,newton_tess_2019,david_four_2019,bouma_cluster_2020,plavchan_planet_2020,martioli_aumicbc_2021}.
\pn\ is one of the smallest planets in this sample ($3.34\pm
0.18\,R_\oplus$), which could be linked to the selection effects
imposed by the spot-induced photometric variability at very young ages
\citep[{\it e.g.},][]{zhou_2021_tois}.  Assuming that these young
planets have masses between $\approx$5$\,M_\oplus$ and 10$\,M_\oplus$,
their large sizes and presumed low densities would be in accord with
the expectation that mini-Neptunes start their lives with large
primordial atmospheres that are shed over the first gigayear
\citep{Owen_Wu_2013,Fulton_et_al_2017,ginzburg_corepowered_2018}.  Our
suggestion of a prograde orbit is consistent with a quiescent
disk-driven migration history, and differs from misalignments that
have been observed for some older
mini-Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
Spectroscopic follow-up observations of the transit remain necessary
however to verify whether the TTV-slope correlation is yielding the
correct orbital geometry.

@@

Ultimately, the main advance from this work is a precise measurement of
the age of \pn.   This measurement was enabled by identifying the
connection of the star to the \cn\ using Gaia kinematics, and then by
using the Gaia HR diagram and TESS stellar rotation periods to verify the
cluster's existence.
Table~\ref{tab:v06} should enable similar cross-matches to already
known planets, as well as forthcoming planets from TESS
\citep{guerrero_tess_2021}.  An alternative
approach for age-dating more planets could be to identify kinematic
associations around known exoplanet host stars using positions and
tangential velocities from Gaia, and to then verify these associations
with ancillary rotation periods and spectroscopy \citep[{\it
e.g.},][]{tofflemire_tess_2021}.

>>>>

Beyond the kinematic and rotational lines of evidence for the youth of
\sn, there is a final consistency check: the abundance of
Li\,\textsc{I} as inferred from the 6708\,\AA\ doublet \citep[{\it
e.g.},][]{soderblom_ages_2014}.  We acquired an iodine-free spectrum
from Keck/HIRES on the night of 2021 March 26 using the standard setup
and reduction techniques of the California Planet Survey
\citep{howard_cps_2010}.  Following the equivalent width measurement
procedure described by \citet{bouma_2021_ngc2516}, we find ${\rm
EW}_{\rm Li} = 233^{+5}_{-7}$\,m\AA.   %Note that this value does not
%correct for the Fe\,\textsc{I} blend at 6707.44\AA. 
Given the effective temperature of \sn\ (Table~\ref{tab:starparams}),
this measurement is in agreement with expectations for a 38\,Myr star
({\it e.g.}, as measured in IC~2602 by
\citealt{randich_gaiaeso_2018}).  It is also larger than any
lithium equivalent widths measured by \citet{berger_identifying_2018}
in their analysis of 1{,}301 Kepler-star spectra.

@@

		\vspace{-0.2cm}	
		\subfloat{
			\includegraphics[width=0.5\textwidth]{f5a.pdf}
			\includegraphics[width=0.5\textwidth]{f5b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The light curve of Kepler\,1627.}
    {\it Top}: 
    The Kepler data span 1{,}437 days (3.9 years), sampled at
    30 minute cadence;  a 100 day segment is shown.  The
    top panel shows the \texttt{PDCSAP} median-subtracted flux in
    units of parts-per-thousand ($\times 10^{-3}$).  The dominant
    signal is induced by starspots.  The stellar
    variability model (orange line) is subtracted below, revealing the
    transits of \pn.  The
    online Figure Set spans the entire 3.9 years of observations.
    {\it Bottom}:
    Phase-folded transit of Kepler 1627Ab with stellar variability
    removed.  Windows over 20 hours ({\it left}) and the entire orbit
    ({\it right}) are shown, and the residual after subtracting the
    transit is in the bottom-most row.  The 2-$\sigma$ model uncertainties
    and the best-fit model are the light purple band and the dark
    purple line.  Gray points are individual flux measurements; black
    points bin these to 20 minute intervals, and have a representative
    1-$\sigma$ error bar in the lower right of each panel.  The
    asymmetric residual during transit appears to be robust to
    different detrending methods.
    \label{fig:lc}
  }
\end{figure*}

@@

The second line of confirmatory evidence comes by analyzing the
individual Kepler transit times. We isolated each of the 144 observed
transits to within $\pm4.5$\,hr of each transit, and fitted each
window with both {\it i)} a local second-order polynomial and transit,
and {\it ii)} a local linear trend.  We let the mid-time of each
transit float, and then calculated the residual between the measured
mid-time and that of a periodic orbit.  This residual, the transit
timing variation (TTV), is plotted in Figure~\ref{fig:obliquity}
against the local linear slope.  A significant correlation of $-0.0564
\pm 0.0098\ {\rm ppt\,day}^{-1}$ is observed.  Fewer than ten Kepler
Objects of Interest have shown this correlation
\citep{holczer_time_2015}, which is most readily interpreted as a TTV
induced by unresolved starspot crossings \citep{mazeh_time_2015}.
This is only possible if the planet transits the primary star, which
excludes a background eclipsing binary scenario, and therefore
confirms that \pn\ is a planet.  It also suggests that the planet's
orbit is prograde.  The latter point assumes that the dominant
photometric variability is induced by dark spots, and not bright
faculae.  Given the observed transition of Sun-like stellar
variability from spot to faculae-dominated regimes between young and
old ages, we expect this assumption to be reasonably secure
\citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.

A third supporting line of evidence for the planetary interpretation
also exists, though we consider it less definitive than the stellar
density and TTV-local slope correlation.  We observed a transit of
\pn\ on the night of 2021 Aug 7 simultaneously with Keck/HIRES and
MuSCAT3.  We scheduled the observations using the ephemeris of
\citet{holczer_transit_2016}.  Although we did not detect the
Rossiter-McLaughlin (RM) anomaly, the multi-band MuSCAT3 light curves
show the transit (Figure~\ref{fig:obliquity}, lower panels).  Fitting
the MuSCAT3 photometry with a model that lets the transit depths vary
across each bandpass, we find {\it griz} depths consistent with the
Kepler depth at 0.6, 0.3, 0.3, and 1.1-$\sigma$ respectively.  The
MuSCAT3 observations also suggest a transit duration $17.4\pm3.6\ {\rm
minutes}$ shorter than the Kepler transits; further photometric
follow-up could help confirm whether the transit duration is indeed
changing.  For our RM analysis, the details are discussed in
Appendix~\ref{app:rm}.  While the velocities are marginally more
consistent with a prograde or polar orbit than a retrograde orbit, the
spot-corrected exposure-to-exposure scatter ($\sigma_{\rm RV}\approx
50$\ms) is larger than the expected velocity anomaly
($\Delta v_{\rm RM}\approx 20\,{\rm m\,s}^{-1}$).  We are therefore
not in a position to claim a spectroscopic detection of the RM effect,
nor to quantify the stellar obliquity.

@@

Kepler\,1627Ab provides a new extremum in the ages of the Kepler
planets, and opens multiple avenues for further study.  Further
observations of spectroscopic transits could yield a stellar obliquity
measurement, which would confirm or refute the prograde orbital
geometry suggested by the TTV-local slope correlation.  Separately,
transit spectroscopy aimed at detecting atmospheric outflows could
yield insight into the evolutionary state of the atmosphere
\citep[{\it
e.g.},][]{ehrenreich_giant_2015,spake_helium_2018,vissapragada_2020}.
Observations aimed at quantifying the amount of high-energy
irradiation incident on the planet would complement these efforts, by
helping to clarify the expected outflow rate \citep[{\it
e.g.},][]{poppenhaeger_2021}.  Finally, a challenging and informative
quantity to measure would be the planet's mass.  This might be
possible through a high-cadence multi-color radial velocity campaign,
or through long-term monitoring of the transit times and durations.
Measured at sufficient precision, the mass, combined with the known
age and size, would yield constraints on both the planet's composition
and its initial entropy \citep{owen_constraining_2020}.

More immediately, the Kepler data may contain additional
information.  For instance, one possible explanation for the transit
asymmetry shown in Figure~\ref{fig:lc} is that of a dusty asymmetric
outflow.  Other possible explanations such as favored longitudes for
starspot crossings seem strained for reasons discussed in
Appendix~\ref{app:gptransit}.  Dusty outflows are theoretically
expected for young mini-Neptunes, and the amplitude of the observed
asymmetry is also roughly consistent with predictions
\citep{wang_dai_2019}.  Beyond the planetary transits,
Appendix~\ref{app:flare} highlights a separate oddity in the Kepler
data, in the arrival time distribution of stellar flares.  We
encourage its exploration by investigators more versed in the topic
than ourselves.

In the context of the transiting planet population, \pn\ is among the
youngest known (Figure~\ref{fig:rp_period_age}).  Comparable systems
with precise ages include K2-33, DS~Tuc, HIP~67522,
TOI~837, the two-planet AU~Mic system, and the four-planet V1298~Tau
system
\citep{Mann_K2_33b_2016,David_et_al_2017,newton_tess_2019,david_four_2019,bouma_cluster_2020,plavchan_planet_2020,martioli_aumicbc_2021}.
\pn\ is one of the smallest planets in this sample ($3.34\pm
0.18\,R_\oplus$), which could be linked to the selection effects
imposed by the spot-induced photometric variability at very young ages
\citep[{\it e.g.},][]{zhou_2021_tois}.  Assuming that these young
planets have masses between $\approx$5$\,M_\oplus$ and 10$\,M_\oplus$,
their large sizes and presumed low densities would be in accord with
the expectation that mini-Neptunes start their lives with large
primordial atmospheres that are shed over the first gigayear
\citep{Owen_Wu_2013,Fulton_et_al_2017,ginzburg_corepowered_2018}.  The
prograde orbit, if confirmed, would be consistent with a quiescent
disk-driven migration history, and would differ from misalignments that
have been observed for some older
hot Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
%Spectroscopic follow-up observations of the transit remain necessary
%however to verify whether the TTV-slope correlation is yielding the
%correct orbital geometry.

@@

Ultimately, the main advance of this work is a precise measurement of
the age of \pn.   This measurement was enabled by identifying the
connection of the star to the \cn\ using Gaia kinematics, and then by
using the Gaia HR diagram and TESS stellar rotation periods to verify the
cluster's existence.
Table~\ref{tab:v06} should enable similar cross-matches to already
known planets, as well as forthcoming planets from TESS
\citep{guerrero_tess_2021}.  An alternative
approach for age-dating more planets could be to identify kinematic
associations around known exoplanet host stars using positions and
tangential velocities from Gaia, and to then verify these associations
with ancillary rotation periods and spectroscopy \citep[{\it
e.g.},][]{tofflemire_tess_2021}.
